Article Title

### 2014-01-24 plos.tex
Rev 41.

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.  The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.

@@

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).

@@

Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.  This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding. Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.  For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry. Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2012}  with our methods can be used to relate functional connectivity to the circuit architecture. Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions. Extending the estimation framework to temporal domain is important for future work.

>>>>

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.  The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.

@@

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}).

@@

Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.  This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding. Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.  For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry. Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture. Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions. Extending the estimation framework to temporal domain is important for future work.
